Using numerical modelling to investigate temperature and wave exposure as drivers

### 2020-06-29 thesis_chapter_2.Rmd
updates to outputs and rmd files for chapters 1,2 and 3. Updated figures to follow and final chpater 2 to follow

## Make first map of whole study area
map_A <- ggplot() +
geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
            size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
  theme(axis.ticks = element_blank()) + # Remove tick marks
  xlab("Latitude") +
  ylab("Longitude")
ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
## Make 2nd map of just the peninsula
map_B <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
  theme(axis.ticks = element_blank(), # Remove tick marks
        axis.text = element_blank(), # Remove lat/ lon numbers
        axis.title = element_blank()) # Remove lat/ lon labels
#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")

@@

## Create combined map
map_combined <- ggplot() +
  # First set the x and y axis limits so we know what the ranges are
  coord_equal(xlim = c(1, 10), ylim = c(1, 10), ratio = 1, expand = TRUE, clip = "on") +
  # Then we place our facetsover one another using the coordinates we created
  annotation_custom(grob_A,
                    xmin = 1, xmax = 10, ymin = 1, ymax = 10) +
  annotation_custom(grob_B,
                    xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
```

@@

```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
### Plots for temperature variables 
positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")

# Annual temp
temp_plot <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
  geom_point(show.legend = FALSE) + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(9, 21) + 
  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
  geom_point(data = temp_ann, aes(x = site, y = max_temp), show.legend = FALSE, shape = 17, size =2) + 
  geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip() + 
  theme_bw() + 
  theme(axis.text.y=element_text(vjust=0.3))

temp_ann_meta <- merge(meta_all, temp_ann, by = "site")
temp_ann_meta$site <- as.character(temp_ann_meta$site)

@@

@@

```{r Temperature variables plot, echo=FALSE, warning=FALSE, eval=TRUE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A. Temperature variables include minimum represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation. Summary of temperature data for sites grouped by region are represented in plot B.'}
ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("hv"), legend = FALSE)
```

@@

```{r wave variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
### Plots for wave variables
#Annual wave plots
#combine with meta data for easier plotting
wave_ann_meta <- merge(meta_all, wave_ann, by = "site")
wave_ann_meta$site <- as.character(wave_ann_meta$site)#make character to ensure no issues arise for site and region names

positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Millers Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
#cols <- c("Betty's Bay" = "red", "Roman Rock"= "red","Kalk Bay"= "red", "A-Frame"= "red", "Miller's Point"= "red", "Baboon #Rock"= "red",  "Batsata Rock"= "red","Bordjies Deep"= "red","Buffelsbaai"= "red","Buffelsbaai South"= "red","Olifantsbos"= #"blue","Scarborough"= "blue", "Soetwater"= "blue", "Kommetjie"= "blue","Hout Bay"= "blue","Oudekraal"= "blue","Sea Point"= #"blue","St Helena Bay"= "blue") #  scale_colour_manual(values = cols) #script to assign colours to specific sites

wave_plot_1 <- ggplot(data = abio_clim, aes(x = site, y = hs_mean)) +
  geom_point(show.legend = FALSE, shape = 17, size = 2) + 
  scale_x_discrete(limits = positions_wave) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = hs_mean - hs_sd, ymax = hs_mean + hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip() + 
  theme_bw() + 
  theme(axis.text.y=element_text(vjust=0.3))

wave_plot_2 <- ggplot(data = abio_clim, aes(x = site, y = tp_mean)) +
  geom_point(show.legend = FALSE, shape = 16, size = 2) + 
  scale_x_discrete(limits = positions_wave) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = tp_mean - tp_sd, ymax = tp_mean + tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip() +
  theme_bw() + 
  theme(axis.text.y=element_text(vjust=0.3))

@@

```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}Mean significant wave height and mean significant peak period across sites and by region. Mean significant wave height across sites is presented in plot A and is represented by solid triangles and standard deviation by the whiskers. Differences in mean significant wave height between regions are presented in plot B. Mean significant peak period across sites is presented in plot C and is represented by solid dots and standard deviation by whiskers. Differences in mean significant peak period between regions are presented in plot D.'}
ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = FALSE, labels = c("A", "B", "C", "D"), align = "hv")
```

@@

#temperature and lamanaria  
LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.62, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina weight") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina length") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina thickness") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

@@

LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe diameter (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey()

LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("No. of digits") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

@@

LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.85, label.y.npc = c('top')) + 
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("Lamina thickness") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe diameter (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("No. of digits") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Thallus mass (kg)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

pos_temp <- c("Cool temperate", "Warm temperate") #custom temperature categories position on x-axis
#temperature plots made seperately for each morphological characteristic
ET1 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  ylab("Frond mass (g)") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

>>>>

## Make first map of whole study area
map_A <- ggplot() +
geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
            size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
  geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
  annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "Cool\nTemperate", x = 18.0, y = -34.0, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
  theme(axis.ticks = element_blank()) + # Remove tick marks
  xlab("Latitude") +
  ylab("Longitude")
#ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")

## Make 2nd map of just the peninsula
map_B <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
  geom_label(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -33.915)) +
  theme(axis.ticks = element_blank(), # Remove tick marks
        axis.text = element_blank(), # Remove lat/ lon numbers
        axis.title = element_blank()) # Remove lat/ lon labels
#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")

@@

## Create combined map
map_combined <- ggplot() +
  # First set the x and y axis limits so we know what the ranges are
  coord_equal(xlim = c(1, 10), ylim = c(1, 10), ratio = 1, expand = TRUE, clip = "on") +
  # Then we place our facetsover one another using the coordinates we created
  annotation_custom(grob_A,
                    xmin = 1, xmax = 10, ymin = 1, ymax = 10) +
  annotation_custom(grob_B,
                    xmin = 6.2, xmax = 10, ymin = 3.8, ymax = 10)
#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8, dpi = 600)
```

@@

```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}

### Plots for temperature variables

#temp_ann$site <- as.character(temp_ann$site)#make character to ensure no issues arise for site and region names
#meta_all$site <- as.character(meta_all$site)#make character to ensure no issues arise for site and region names

#temp_ann_meta_v2 <- left_join(temp_ann, meta_all, by = "site", copy = FALSE, keep = TRUE)

region <- c("Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Western", "Eastern","Western", "Eastern", "Western", "Western", "Eastern", "Western", "Western", "Western", "Western", "Western")

temp_ann$region <- region
temp_ann$region <- as.factor(temp_ann$region)

temp_ann_meta <- merge(meta_all, temp_ann)

positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")

# Annual temp
temp_plot <- ggplot(data = temp_ann_meta, aes(x = site, y = mean_temp)) +
  geom_point(show.legend = FALSE, aes(fill = region)) +
  scale_x_discrete(limits = positions_temp) +
  ylim(9, 21) +
  geom_errorbar(data = temp_ann, aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) +
  geom_point(data = temp_ann, aes(x = site, y = max_temp), show.legend = FALSE, shape = 17, size =2) +
  geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) +
  geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) +
  geom_point(data = temp_ann, aes(x = site, y = mean_temp, colour = region), show.legend = FALSE, size =2) +
  xlab("") +
  ylab("") +
  theme(legend.position = "none") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.y=element_text(vjust=0.3))

#ggsave("fig_3A.jpeg", plot = temp_plot, dpi = 300, device = "jpeg" , path = "figures/")

@@

#ggsave("fig_3B", plot = temp_boxplot, dpi = 300, device = "jpeg", path = "figures/frontiers_fig_prep/")

@@

```{r Temperature variables plot, eval=TRUE, echo=FALSE, fig.cap='\\label{fig:Temperature variables plot}Annual temperature variables at the collections sites around the Cape Peninsula are represented by plot A, with sites in order from North to South. Temperature variables include minimum temperature represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation. Summary of temperature data for sites grouped by region and represented in plot B.', fig.height=5, fig.width=8, warning=FALSE}

ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("hv"), legend = FALSE)

#ggsave("fig_3", plot = temp_boxplot, dpi = 300, device = "jpeg", path = "figures/")

@@

```{r wave variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
### Plots for wave variables
#Annual wave plots
#combine with meta data for easier plotting

wave_ann$site <- as.character(wave_ann$site)#make character to ensure no issues arise for site and region names
wave_ann_meta <- merge(meta_all, wave_ann, by = "site")

wave_ann_meta$site <- as.character(wave_ann_meta$site)#make character to ensure no issues arise for site and region names
wave_ann_meta <- wave_ann_meta %>%
  mutate(site = replace(site, site == "Miller's Point C", "Miller's Point"))

#wave_ann_meta$site <- as.character(wave_ann_meta$site)#make character to ensure no issues arise for site and region names

positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
#cols <- c("Betty's Bay" = "red", "Roman Rock"= "red","Kalk Bay"= "red", "A-Frame"= "red", "Miller's Point"= "red", "Baboon #Rock"= "red",  "Batsata Rock"= "red","Bordjies Deep"= "red","Buffelsbaai"= "red","Buffelsbaai South"= "red","Olifantsbos"= #"blue","Scarborough"= "blue", "Soetwater"= "blue", "Kommetjie"= "blue","Hout Bay"= "blue","Oudekraal"= "blue","Sea Point"= #"blue","St Helena Bay"= "blue") #  scale_colour_manual(values = cols) #script to assign colours to specific sites

wave_plot_1 <- ggplot(data = wave_ann_meta, aes(x = site, y = hs_mean)) +
  geom_point(show.legend = FALSE, shape = 17, size = 2, aes(colour = region)) + 
  scale_x_discrete(limits = positions_wave) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = hs_mean - hs_sd, ymax = hs_mean + hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip() + 
  theme_bw() + 
  theme(axis.text.y=element_text(vjust=0.3))

wave_plot_2 <- ggplot(data = wave_ann_meta, aes(x = site, y = tp_mean)) +
  geom_point(show.legend = FALSE, shape = 16, size = 2, aes(colour = region)) + 
  scale_x_discrete(limits = positions_wave) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = tp_mean - tp_sd, ymax = tp_mean + tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip() +
  theme_bw() + 
  theme(axis.text.y=element_text(vjust=0.3))

@@

```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}Mean significant wave height and mean significant peak period across sites and by region. Mean significant wave height across sites is presented in plot A and is represented by solid triangles and standard deviation by the whiskers. Differences in mean significant wave height between regions are presented in plot B. Mean significant peak period across sites is presented in plot C and is represented by solid dots and standard deviation by whiskers. Differences in mean significant peak period between regions are presented in plot D.'}
ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = FALSE, labels = c("A", "B", "C", "D"), align = "hv")

ggsave(plot = last_plot(), filename = "Fig_3", path = "figures/chapter_2/", device = "jpeg", units = "cm", width = 7, height = 7)
```

@@

#temperature and lamanaria  
LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.62, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina weight (kg)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina length (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina thickness (mm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe mass (kg)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

@@

LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe diameter (mm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey()

LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("No. of digits (count)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

@@

LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.85, label.y.npc = c('top')) + 
  ylab("Lamina weight (kg)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("Lamina thickness (mm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe mass (kg)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe diameter (mm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("No. of digits (count)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Thallus mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

pos_temp <- c("Cool temperate", "Warm temperate") #custom temperature categories position on x-axis
#temperature plots made seperately for each morphological characteristic
ET1 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  ylab("Frond mass (kg)") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)
